---
title: "Introduction to smokingMouse"
author: 
  - name: Daianna Gonzalez-Padilla
    affiliation:
    - Lieber Institute for Brain Development (LIBD)
    email: daianna2121@outlook.es
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('smokingMouse')`"
vignette: >
  %\VignetteIndexEntry{Introduction to smokingMouse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```


```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## Track time spent on making the vignette
startTime <- Sys.time()

## Bib setup
library("RefManageR")

## Write bibliography information
bib <- c(
    R = citation(),
    AnnotationHub = citation("AnnotationHub")[1],
    cowplot = citation("cowplot")[1],
    edgeR = citation("edgeR")[1],
    ExperimentHub = citation("ExperimentHub")[1],
    BiocStyle = citation("BiocStyle")[1],
    knitr = citation("knitr")[1],
    RefManageR = citation("RefManageR")[1],
    rmarkdown = citation("rmarkdown")[1],
    sessioninfo = citation("sessioninfo")[1],
    testthat = citation("testthat")[1],
    smokingMouse = citation("smokingMouse")[1],
    here = citation("here")[1],
    jaffelab = citation("jaffelab")[1],
    SummarizedExperiment = citation("SummarizedExperiment")[1],
    ggplot2 = citation("ggplot2")[1],
    rlang = citation("rlang")[1],
    scater = citation("scater")[1],
    ggrepel = citation("ggrepel")[1],
    gridExtra = citation("gridExtra")[1],
    variancePartition = citation("variancePartition")[1],
    stats = citation("stats")[1],
    limma = citation("limma")[1],
    VennDiagram = citation("VennDiagram")[1],
    biomartr = citation("biomartr")[1],
    Hmisc = citation("Hmisc")[1],
    R.utils = citation("R.utils")[1],
    clusterProfiler = citation("clusterProfiler")[1],
    org.Mm.eg.db = citation("org.Mm.eg.db")[1],
    pheatmap = citation("pheatmap")[1],
    GenomicRanges = citation("GenomicRanges")[1],
    rtracklayer = citation("rtracklayer")[1],
    AcidGenomes = citation("AcidGenomes")[1]
)
```

# Welcome

Welcome to the `smokingMouse` project.

You can find the analysis code and the data generation in [here](https://github.com/LieberInstitute/smokingMouse_Indirects).

## Overview
This project consisted of a differential expression analysis involving 4 data types: genes, exons, transcripts and junctions. The main goal of this study was to explore the effects of smoking and nicotine exposures on the developing brain of mice pups. As secondary objectives, this work evaluated the affected genes by each substance on adult brain in order to compare pup and adult results, and the effects of smoking on adult blood and brain to search for overlapping biomarkers in both tissues.


## Study design
36 pregnant mice and 35 not pregnant female adults were administered nicotine (n=12), exposed to cigarette smoke (n=24) or controls (n=35) and RNA sequencing experiments were performed on frontal cortices of all the resultant 137 P0 pups and on frontal cortices (n=47) and blood (n=24) from the 71 adults, totaling 208 samples. Of the total pup samples, 19 were born to mice that were administered nicotine, 46 to mice exposed to smoking and the remaining 72 to control mice.


# Basics

## Install `smokingMouse`

`R` is an open-source statistical environment which can be easily modified to enhance its functionality via packages. `r Biocpkg("smokingMouse")` is a `R` package available via the [Bioconductor](http://bioconductor.org) repository for packages. `R` can be installed on any operating system from [CRAN](https://cran.r-project.org/) after which you can install `r Biocpkg("smokingMouse")` by using the following commands in your `R` session:

```{r "install", eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
      install.packages("BiocManager")
  }

BiocManager::install("smokingMouse")

## Check that you have a valid Bioconductor installation
BiocManager::valid()
```

## Required knowledge

`r Biocpkg("smokingMouse")` is based on many other packages and in particular in those that have implemented the infrastructure needed for dealing with RNA-seq data and differential expression. That is, packages like `r Biocpkg("SummarizedExperiment")` and `r Biocpkg("limma")`.

If you are asking yourself the question "Where do I start using Bioconductor?" you might be interested in [this blog post](http://lcolladotor.github.io/2014/10/16/startbioc/#.VkOKbq6rRuU).


## Asking for help

As package developers, we try to explain clearly how to use our packages and in which order to use the functions. But `R` and `Bioconductor` have a steep learning curve so it is critical to learn where to ask for help. The blog post quoted above mentions some but we would like to highlight the [Bioconductor support site](https://support.bioconductor.org/) as the main resource for getting help: remember to use the `smokingMouse` tag and check [the older posts](https://support.bioconductor.org/tag/smokingMouse/). Other alternatives are available such as creating GitHub issues and tweeting. However, please note that if you want to receive help you should adhere to the [posting guidelines](http://www.bioconductor.org/help/support/posting-guide/). It is particularly critical that you provide a small reproducible example and your session information so package developers can track down the source of the error.

## Citing `smokingMouse`

We hope that `r Biocpkg("smokingMouse")` will be useful for your research. Please use the following information to cite the package and the overall approach. Thank you!

```{r "citation"}
## Citation info
citation("smokingMouse")
```

# Quick start to using `smokingMouse`

To get started, please load the `r Biocpkg('spatialLIBD')` package. 

```{r "start", message=FALSE}
library("smokingMouse")
```

Edit this as you see fit =)

Here is an example of you can cite your package inside the vignette:

* `r Biocpkg("smokingMouse")` `r Citep(bib[["smokingMouse"]])`











# Reproducibility

The `r Biocpkg("smokingMouse")` package `r Citep(bib[["smokingMouse"]])` was made possible thanks to:

* R `r Citep(bib[["R"]])`
* `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
* `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])`
* `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`
* `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])`
* `r CRANpkg("sessioninfo")` `r Citep(bib[["sessioninfo"]])`
* `r CRANpkg("testthat")` `r Citep(bib[["testthat"]])`

This package was developed using `r BiocStyle::Biocpkg("biocthis")`.


Code for creating the vignette

```{r createVignette, eval=FALSE}
## Create the vignette
library("rmarkdown")
system.time(render("smokingMouse.Rmd", "BiocStyle::html_document"))

## Extract the R code
library("knitr")
knit("smokingMouse.Rmd", tangle = TRUE)
```

Date the vignette was generated.

```{r reproduce1, echo=FALSE}
## Date the vignette was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r reproduce2, echo=FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```



# Bibliography

This vignette was generated using `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])` and `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])` running behind the scenes.

Citations made with `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`.

```{r vignetteBiblio, results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```
